---
title: "Assignment3"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Create a R Markdown powered Web Application w/ R & Shiny:
### Share it with github

1. **README (10 pt) file in top level directory.** 
+ The readme will contain your name
2. **LICENSE** An MIT License (10 pt)**
3. **original** Directory containing assignment named original
4. **original/index.Rmd** File containing the text here without editing
5. **index.Rmd**  R markup file containing your completed homework
6. **data** A directory containing data we will use
7. **data/gene_id2gene.tsv.gz** A file containing gene_id and gene mapping
8. **data/cohort.tpms.csv.gz** A file containing expression levels for a set of samples


# Question 1

## Question 1 Background
We will use example data from R, the iris dataset.  For the assignment, I'll ask for you to complete things from the diamonds or iris dataset. These are preloaded datasets

data(diamonds)
summary(diamonds)
  
```{r}
  library('rmarkdown')
  library("ggplot2")
  library("plotly")
  library("DT")
  diamonds<-diamonds
  iris<-iris
  summary(iris)
```


### Question 1 Background: Analysis of our iris dataset
Example functions and how to use them to calculate some stats manually.
Remember one can look these up online.

```{r}
# Calculates the mean values for the aggregates given by the Species column in the iris data set.
 iris_mean <- aggregate(iris[,1:4], by=list(Species=iris$Species), FUN=mean)
# Calculates the standard deviations for the aggregates given by the Species column in the iris data set.
 iris_sd <- aggregate(iris[,1:4], by=list(Species=iris$Species), FUN=sd)
 convertDF <- function(df=df, mycolnames=c("Species", "Values", "Samples")) { myfactor <- rep(colnames(df)[-1], each=length(df[,1]));
# Defines function to convert data frames into ggplot2-friendly format.
mydata <- as.vector(as.matrix(df[,-1])); df <- data.frame(df[,1], mydata, myfactor); colnames(df) <- mycolnames; return(df) }
# Converts iris_mean.
 df_mean <- convertDF(iris_mean, mycolnames=c("Species", "Values", "Samples"))
# Converts iris_sd.
 df_sd <- convertDF(iris_sd, mycolnames=c("Species", "Values", "Samples"))
# Define standard deviation limits.
 limits <- aes(ymax = df_mean[,2] + df_sd[,2], ymin=df_mean[,2] - df_sd[,2])
```

## **Question 1** (20 Points): 
1. Please put the Mean Petal Length into mean_petal_length variable
 + mean_petal_length <- mean(iris$'petal.length') 3.758
2. Please put the Max Petal Length max_petal_length variable
+ max_petal_length <- max(iris$'petal.length') 6.9
3. How many vectors (columns) are there in Iris and what are there names?
+ There are 5 Columns. Sepal.Length, Sepal.width, Petal.length, petal.width
ncol(iris)
names(iris)
```{r}
mean_petal_length <- mean(iris$'Petal.Length')

mean_petal_length

max_petal_length <- max(iris$'Petal.Length')

max_petal_length

ncol(iris)

names(iris)

```


# Question 2





### Question 2 Background

### Question 2 Background: Examples with *ggplot2*

  Ploting two vectors (columns) in form of a scatter plot against each other and places into a png image called *twovectors.png* in the images directory.  Make sure you have made your directory
  
```{r}
dir.create('images')
ggplot(iris, aes(Sepal.Length, Sepal.Width)) + geom_point()
png('images/twovectors.png')
ggplot(iris, aes(Sepal.Length, Sepal.Width)) + geom_point()
dev.off()
```

### Question 2 Background Example A: Plots larger dots and colors them with default color scheme.

```{r example1}
example1 <- ggplot(iris, aes(Sepal.Length, Sepal.Width)) + geom_point(aes(color = Species), size=4)
return(example1)
```

### Question 2 Background Example B: Colors dots with custom scheme.
```{r example2}
example2 <- ggplot(iris, aes(Sepal.Length, Sepal.Width)) + geom_point(aes(color = Species), size=4) + ylim(2,4) + xlim(4,8) + scale_color_manual(values=rainbow(10))
return(example2)
```

### Question 2 Background Example C: Adds a regression line to the plot.
```{r example3}
example3 <- ggplot(iris, aes(Sepal.Length, Sepal.Width)) + geom_point() + stat_smooth(method="lm", se=FALSE) + ylim(2,4) + xlim(4,8) + scale_color_manual(values=rainbow(10))
return(example3)
```
### Question 2 Background Example D: Plots three line plots, one for each sample in Species column.
```{r example4}
example4 <- ggplot(iris, aes(Sepal.Length, Sepal.Width)) + geom_line(aes(color=Species), size=1) + facet_wrap(~Species, ncol=1)
return(example4)
```

### Question 2 Background Example E: Plots with q plot
Qplot is another way to do ggplot2, but a bit simpler
```{r example5}
example5<-qplot(Sepal.Length, Petal.Length, data = iris, color = Species, size = Petal.Width)
return(example5)
```


### Question 2 Background Example F
#### Same with ggplot & _Plotly_

```{r example6}
a<-ggplot(iris, aes(x = Sepal.Length, y = Petal.Length, color=Species)) + geom_point() + facet_wrap(~Species) + ggtitle("Plot of sepal length vs. petal length")
example6<-ggplotly(a)
return(example6)
```

## **Question 2** (20 Points): 
Using **diamonds** instead of **iris** model data place graph of carat (x-axis) vs. price (y-axis), color by clarity, facet by cut with a title (Based on Example 2F).

```{r Answer to Question 2 carat vs. price}
diamonds<-diamonds
b<-ggplot(diamonds, aes(carat, price, color=clarity)) + geom_point() + facet_wrap(~cut) + ggtitle("Plot of Carat vs. Price")
return(b)

```


# **Question 3** 

### Question 3 Background Building a Widget
*Lets create our inputs (remember to make these code)*

<pre>
inputPanel(
  selectInput("var1", label = "Choose Xaxis:", 
     choices = c("Sepal.Length","Petal.Length"), selected = "Petal.Length"
     ),
  selectInput("var2",  label = "Choose Yaxis:", 
     choices = c("Sepal.Width","Petal.Width"), selected = "Petal.Width"
     )
)
</pre>
*Lets create our outputs (remember to make these code)*

<pre>
renderPlot({
  x_axis <- input$var1
  y_axis <- input$var2
  species <- "Species"
  example13<-ggplot(iris, aes_string(x = x_axis, y = y_axis, color=species)) +
      geom_point() + facet_wrap(~Species) + ggtitle("Plot of sepal length vs. petal length")
  example13
})
</pre>

### Example 3A
```{r echo = FALSE}
inputPanel(
  selectInput("var1", label = "Choose Xaxis:", choices = c("Sepal.Length","Petal.Length"), selected = "Petal.Length"),
  selectInput("var2",  label = "Choose Yaxis:", choices = c("Sepal.Width","Petal.Width"), selected = "Petal.Width")
)
```


```{r  echo = FALSE}
library(plotly)
renderPlotly({
  x_axis <- input$var1
  y_axis <- input$var2
  species <- "Species"
   a<-ggplot(iris, aes_string(x = x_axis, y = y_axis, color=species)) + geom_point() + facet_wrap(~Species) + ggtitle("Plot of sepal length vs. petal length")
   ggplotly(a)
}) 
```

## **Question 3** (20 Points): 
Make a similar shiny app, based on 3A with **diamonds** modelset allowing changing of axis.

```{r echo = FALSE}
inputPanel(
  selectInput("var3", label = "Choose Xaxis:", choices = c("carat","x", "y"), selected = "carat"),
  selectInput("var4",  label = "Choose Yaxis:", choices = c("depth","table", "price"), selected = "price")
)

```

```{r echo = FALSE}
dia<-diamonds
library(plotly)
renderPlotly({
  x_axis <- input$var3
  y_axis <- input$var4
  "Clarity" <- "Clarity"
  ve<-ggplot(dia, aes_string(x = x_axis, y = y_axis, color="clarity")) + geom_point() + facet_wrap(~clarity) + ggtitle("Plot of Carat vs. Price")
   ggplotly(ve) 
   })
```






<!-- End of Assignment -->

<!-- The text below is just to reformat the page, using css styling -->
```{r results="asis", echo = FALSE}
cat("
<style>
h1 {
    background-color: #990000;
    color:#FFF;
    text-align:center;
}
h2 {
   background-color:#f3d6d6;
   text-align:center;
   color:#333
}
h3 {
   margin-left:-5px;
   text-align:center;
   color:#333
}

</style>
")
```


